cardinalR: Generating interesting high-dimensional data structures

### 2025-09-20 paper-cardinalR.Rmd
Updated multicluster section

The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics. Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`) for all the clusters. Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`). The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures. An optional argument, `is_bkg`, adds background noise drawn from a multivariate normal distribution centered on the dataset’s overall mean with standard deviations matching the observed spread.

@@

The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures. These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures. If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space. Users can specify how these noise dimensions are generated (e.g., Gaussian, uniform), offering control over the embedding process. All shape generators allow the user to define the number of points and dimensions, and most include additional arguments to customize specific characteristics of the structure.

>>>>

The main function of the package is `gen_multicluster()`, which generates datasets consisting of multiple clusters with user-specified characteristics. Users can control the number of clusters (`k`), the number of points in each cluster (`n`), and the dimensionality of the space (`p`) for all the clusters. Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (`shape`), can be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (`loc`). The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures. An optional argument, `is_bkg`, adds background noise drawn from a multivariate normal distribution centered on the dataset’s overall mean with standard deviations matching the observed spread. Extra arguments (`...`) can be passed to cluster generators, allowing further control over per-cluster characteristics such as noise structure.

@@

The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures. These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures. If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space. Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (`noise_fun`), offering control over the embedding process. All shape generators allow the user to define the number of points (`n`) and dimensions (`p`), and most include additional arguments to customize specific characteristics of the structure.

### 2025-09-20 paper-cardinalR.tex
Updated multicluster section

The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics. Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}) for all the clusters. Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube), be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}). The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures. An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread.

@@

The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures. These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures. If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space. Users can specify how these noise dimensions are generated (e.g., Gaussian, uniform), offering control over the embedding process. All shape generators allow the user to define the number of points and dimensions, and most include additional arguments to customize specific characteristics of the structure.

>>>>

The main function of the package is \texttt{gen\_multicluster()}, which generates datasets consisting of multiple clusters with user-specified characteristics. Users can control the number of clusters (\texttt{k}), the number of points in each cluster (\texttt{n}), and the dimensionality of the space (\texttt{p}) for all the clusters. Each cluster can take on a different geometric shape (e.g., Gaussian, cone, uniform cube) by specifying the corresponding generator function (\texttt{shape}), can be scaled to adjust its spread, rotated in specified planes by given angles, and positioned at defined centroids (\texttt{loc}). The function ensures flexibility in cluster location and orientation, allowing users to simulate complex high-dimensional structures. An optional argument, \texttt{is\_bkg}, adds background noise drawn from a multivariate normal distribution centered on the dataset's overall mean with standard deviations matching the observed spread. Extra arguments (\texttt{...}) can be passed to cluster generators, allowing further control over per-cluster characteristics such as noise structure.

@@

The shape generators form the foundation of the package, providing a collection of functions to create synthetic data structures based on simple, well-defined geometric structures. These include fundamental shapes such as cones, pyramids, spheres, grids, and branching structures. If a shape is not inherently defined in more than three dimensions, additional noise dimensions can be added to embed the structure into higher-dimensional space. Users can specify how these noise dimensions are generated (e.g., Gaussian, wavy) (\texttt{noise\_fun}), offering control over the embedding process. All shape generators allow the user to define the number of points (\texttt{n}) and dimensions (\texttt{p}), and most include additional arguments to customize specific characteristics of the structure.
